A decentralized network support middleware for MMOGs

### 2009-09-05 textos/PlanoDoutorado/planodoc_lugano.tex

Another approach is the \textbf{peer-to-peer architecture}, where the simulation is divided among the computers involved. There can be a system without any server, where the peers, which are the machines of the players, come into some kind of agreement for the various steps of the simulation, calculating the game state resulting of each action of the players. Regarding scalability, this approach is not optimal, since ensuring such "agreement"{} is costly in terms of message exchange \cite{lamport1982bgp}. Even if one of the peers is elected to decide the course of the simulation, still persists the problem of all peers needing to exchange messages with one another. Having $n$ peers, there is a complexity of O({$n^2$}) message exchanges for each step of the simulation. Clearly, this approach is not as scalable as should be a system to support a massively multiplayer online game. Moreover, it would be necessary to distribute the storage and retrieval of the states of the game.

Some studies -- such as \cite{schiele2007rpp, rieche2007ppb, hampel2006ppa, elrhalibi2005abm, iimura2004zfg, knutsson2004pps} -- have been done to make games on peer-to-peer networks more scalable, . For example, to reduce traffic between the peers, each peer can send state updates only to whom the state changes are relevant. To achieve this goal, \cite{schiele2007rpp} suggests the division of the virtual environment into regions. The purpose of this is for each region to act as an independent set of smaller scale. Players whose avatars were in a given region could form a small peer-to-peer group and their own computers would decide the outcome of the players' actions in that area of the virtual environment. When an avatar moved from one region to another, it would leave the first group and join the group of its new region. Since the environment is considered contiguous, i.e. the avatars could move freely across adjacent regions, some of the peers in a group would keep connections to peers from groups of neighboring regions and, through these connection between adjacent groups, the necessary information for transfering a player could be obtained.

%TODO: por mais refs aqui após o ", como"
According to the work of \cite{schiele2007rpp}, each region has a coordinator. The coordinator is elected among the peers which were present and becomes responsible, only, to decide to whom each state update is relevant -- it does \emph{not} intermediate message exchanges between peers, whose communication is direct. However, this approach is based on the fact that the coordinator is reliable, which can not be guaranteed, since the software used by the player may have been modified to behave incorrectly in order to provide invalid benefits -- what is also known as \emph{cheating}. Another approach would be to elect multiple coordinators per region, but that would involve implementation of a voting mechanism, and depend on the availability of a certain number of peers to manage the region. Moreover, the communication load remains quadratic (in the worst case), since every peer must send its state to several other peers.

There are also proposals for \textbf{hybrid architectures} \cite{vilanova2008psm, chen2006pch}, using servers and peer-to-peer networks at the same time. According to these works, peers and server share the simulation of the game world. In \cite{vilanova2008psm}, the game world is divided into social space and action spaces. The first is aimed at social interactions, such as chat, exchange of in-game items, forming groups, etc., which are actions that do not impose a heavy burden on the server. However, if the players want, for example, fight -- what is the main objective in most MMOG games -- they have to request to the server to create an action space within which the interaction is faster and with little tolerance for communication delays. This space is an \emph{instance} of the game, isolated from the rest of the world, within which a limited number of players form a small peer-to-peer group, being themselves responsible for simulating the game and for updating his companions on the state of the game. To resolve inconsistencies in the simulation due to lack of message ordering, one player is chosen as a \emph{super-peer} and is responsible for ordering the events that are "heavily coupled", whose order can significantly interfere in the outcome of the game. This approach presents some problems, such as the fact that there cannot be a fight, for example, between a large number of players. Moreover, nothing guarantees that the peer chosen to be the super-peer is trustable. If it belongs to a dishonest player, it may order the events in an order such that its player is benefited. In addition, the problem of overloading the bandwidth of peers is addressed by simply setting a limit to the number of participants in each action group, when in fact it is desirable that in an MMOGs there is a lot of players interacting with each other, not only in social circumstances, but also in more dynamic situations, like fast-paced fights.

In \cite{chen2006pch}, the virtual environment is divided into regions, each managed by a peer, which acts as a sub-server. Each one of the other participants in the region sends state updates to this sub-server, which forwards the messages to the other peers in the region to which the state changes may be relevant. To provide fault tolerance, the authors suggest the use of peer clusters, consisting of the region's sub-server and backup peers. These backup peers receive the players' actions from the region manager and process them, creating a backup of the game state. In addition to providing robustness to the system in case the manager crashes, these backup nodes may also detect faults in the simulation and perform recovery procedures. However, this approach does not solve the problem of the peer elected to be the region manager not being reliable: to prevent the backup nodes from detecting an invalid simulation, the region manager may send invalid actions to them, instead of forwarding the players' real actions. Finally, if a region manager gets overloaded, it simply returns to the server part of the load assigned to it, keeping a dependency on a centralized infrastructure.

Another hybrid architecture is that of FreeMMG \cite{cecin2004fsa}. In it, the players are organized in a peer-to-peer manner in every region of the virtual environment and the server mediates the communication between different regions. Again, there is the problem of security: as the peers within a region control the simulation that occurs  there, they can also subvert it. The author proposes a probabilistic approach, using a randomly selected peer to verify the simulation in that region. It is expected that, if the peers in a given region try to subvert the game there, at least the randomly inserted node detect the invalid actions and report to the server. However, besides not completely guaranteeing security, the problem persists there may be many peers communicating with many, which may compromise the quality of the game.

@@

It is intended, during the doctoral course, to research and develop new solutions to these problems, seeking those that are most appropriate in a context of massively multiplayer online games. With these issues resolved, it may be provided to MMOGs availability and fault tolerance -- due to the replication of server nodes in the same region -- and scalability, while maintaining the communication delay within an acceptable range for the game and the a lower financial cost, compared to traditional MMOG servers.

@@

\item Create a middleware containing the desired services, along with an high-level API to make them accessible to the game developer.

>>>>

Another approach is the \textbf{peer-to-peer architecture}, where the simulation is divided among the computers involved. There can be a system without any server, where the peers, which are the machines of the players, come into some kind of agreement for the various steps of the simulation, calculating the game state resulting of each action of the participants. Regarding scalability, this approach is not optimal, since ensuring such "agreement"{} is costly in terms of message exchange \cite{lamport1982bgp}. Even if one of the peers is elected to decide the course of the simulation, still persists the problem of all peers potentially needing to exchange messages with one another. Having $n$ peers, there is a complexity of O({$n^2$}) message exchanges for each step of the simulation, in the worst case. Clearly, this approach is not as scalable as should be a system to support a massively multiplayer online game. Moreover, it would be necessary to distribute the storage and retrieval of the states of the game.

Some studies -- such as \cite{schiele2007rpp, rieche2007ppb, hampel2006ppa, elrhalibi2005abm, iimura2004zfg, knutsson2004pps} -- have been done to make games on peer-to-peer networks more scalable. For example, to reduce traffic between the peers, each peer can send state updates only to whom the state changes are relevant. To achieve this goal, \cite{schiele2007rpp} suggests the division of the virtual environment into regions. The purpose of this is for each region to act as an independent set of smaller scale. Players whose avatars were in a given region could form a small peer-to-peer group and their own computers would decide the outcome of the players' actions in that area of the virtual environment. When an avatar moved from one region to another, it would leave the first group and join the group of its new region. Since the environment is considered contiguous, i.e. the avatars could move freely across adjacent regions, some of the peers in a group would keep connections to peers from groups of neighboring regions and, through these connections between adjacent groups, the necessary information for transferring a player could be obtained.

%TODO: por mais refs aqui após o ", como"
Still according to the work of \cite{schiele2007rpp}, each region has a coordinator. The coordinator is elected among the peers which were present in the region and becomes responsible, only, to decide to whom each state update is relevant -- it does \emph{not} intermediate message exchanges between peers, whose communication is direct. However, this approach is based on the fact that the coordinator is reliable, which can not be guaranteed, since the software used by the player may have been modified to behave incorrectly in order to provide invalid benefits -- what is also known as \emph{cheating}. Another approach would be to elect multiple coordinators per region, but that would involve the implementation of a voting mechanism, and depend on the availability of a certain number of peers to manage the region. Moreover, the communication load remains quadratic (in the worst case), since every peer must send its state to several other peers.

There are also proposals for \textbf{hybrid architectures} \cite{vilanova2008psm, chen2006pch}, using servers and peer-to-peer networks at the same time. According to these works, peers and server share the simulation of the game world. In \cite{vilanova2008psm}, the game world is divided into social space and action spaces. The first is aimed at social interactions, such as chat, exchange of in-game items, forming groups, etc., which are actions that do not impose a heavy burden on the server. However, if the players want, for example, fight -- what is the main objective in most MMOGs -- they have to request to the server the creation of an action space within which the interaction is faster and with little tolerance for communication delays. This space is an \emph{instance} of the game, isolated from the rest of the game world, within which a limited number of players form a small peer-to-peer group, being themselves responsible for performing the simulation and for updating his companions on the game state. To resolve inconsistencies in the simulation due to lack of message ordering, one player is chosen as a \emph{super-peer} and is responsible for ordering the events that are "heavily coupled", whose order can significantly interfere in the outcome of the game. This approach presents some problems, such as the fact that there cannot be a fight, for example, between a large number of players. Moreover, nothing guarantees that the peer chosen to be the super-peer is trustable. If it belongs to a dishonest player, it may put the events in an order that provides an invalid advantage over the other players. In addition, the problem of overloading the bandwidth of peers is addressed by simply setting a limit to the number of participants in each action group, when in fact it is desirable that in an MMOGs there is a lot of players interacting with each other, not only in social circumstances, but also in more dynamic situations, like fast-paced fights.

In \cite{chen2006pch}, the virtual environment is divided into regions, each managed by a peer, which acts as a sub-server. Each one of the other participants in the region sends state updates to this sub-server, which forwards the messages to the other peers in the region to which the state changes may be relevant. To provide fault tolerance, the authors suggest the use of peer clusters, consisting of the region's sub-server and backup peers. These backup peers receive the players' actions from the region manager and process them, creating a backup of the game state. In addition to providing robustness to the system in case the manager crashes, these backup nodes may also detect faults in the simulation and perform recovery procedures. However, this approach does not solve the problem of the peer elected to be the region manager 	not being reliable: to prevent the backup nodes from detecting an invalid simulation, the region manager may send invalid actions to them, instead of forwarding the players' real actions. Finally, if a region manager gets overloaded, it may simply return to the server part of the load assigned to it, keeping a dependence on a centralized infrastructure.

Another hybrid architecture is that of FreeMMG \cite{cecin2004fsa}. In it, the players are organized in a peer-to-peer manner in every region of the virtual environment and the server mediates the communication between different regions. Again, there is the problem of security: as the peers within a region control the simulation that occurs  there, they can also subvert it. The author proposes a probabilistic approach, using a randomly selected peer to verify the simulation in that region. It is expected that, if the peers in a given region try to subvert the game there, at least the randomly inserted node detects the invalid actions and reports to the server. However, besides not completely guaranteeing security, persists the problem that there may be many peers communicating with many, which may compromise the quality of the game.

@@

It is intended, during the doctoral course, to research and develop new solutions to these problems, seeking those that are most appropriate in a context of massively multiplayer online games. With these issues resolved, it may be provided to MMOGs availability and fault tolerance -- due to the replication of server nodes in the same region -- and scalability, while maintaining the communication delay within an acceptable range for the game, all this for a low cost for the player, compared to traditional MMOG servers.

@@

\item Create a middleware containing the desired services, along with a high-level API to make them accessible to the game developer.
